A Loop Quantm Gravity Inspired Action for the Bosonic String and Emergent Dimensions at Large Scales

### 2025-07-05 AoP/response-to-referee.tex

\emph{For example, in $D=2$ with $\mu=1$, $a=-4.24$ gives a relatively light spectrum, while $D=3$ with $\mu=5$ yields $a=-34.27$, producing much heavier states. This demonstrates how LQG's discreteness parameter $\mu$ and the background dimensionality jointly control the departure from standard string theory's massless spectrum, with stronger quantum geometric effects pushing all states away from the massless limit expected in conventional string theory.}

\subsection{Section 6 - Part (a): Afshordi Model Connection}

\RC The link to Afshordi's model is interesting, but the constraints on $m_\mu$ (Eq. 6.3) feel ad hoc, a better and deeper connection to LQG's discreteness would strengthen this section.

\AR We acknowledge the reviewer's concern that the constraints on $m_\mu$ (Eq. 6.3) appear ad hoc. Upon reflection, we must be honest that a rigorous connection between Afshordi's mass parameters and LQG's fundamental discreteness has not yet been established. We have added the following two paragraphs to the text right after ``... making it possible to predict in which scales the other dimensions should start to open up.''

\emph{Our constraint equation (6.3) represents an attempt to bridge Afshordi's phenomenological emergent dimension model with LQG's general prediction of modified physics at the Planck scale. However, the specific form of these constraints and their derivation from first-principles LQG remains an open problem. While both frameworks involve discreteness - Afshordi's through hierarchical mass parameters and LQG through quantized geometric operators - establishing a concrete mathematical connection requires significantly more work.}

\emph{This represents an important limitation of our current approach and highlights the need for future research to either derive such constraints from canonical LQG or develop alternative frameworks that more naturally incorporate both emergent dimensions and quantum geometric discreteness. Similar caveats also apply to our suggestion that (6.3) should imply that spacetime is $(2 + 1)$-dimensional.}

@@

\subsection{Section 6 - Part (b): Fundamental Spacetime Dimensionality}

\RC The suggestion of (2+1)-dimensional fundamental spacetime (Sec. 6) is speculative and should be better addressed, more evidence or references to support this claim are needed.

\AR We hope that the last sentence of the second paragraph in the previous point address this point.

@@

\subsection{Section 7 - Part (a): Holographic Dimension Interpretation}

\RC The physical interpretation of the emergent holographic dimension $X^r$ as a "scaling dimension" needs clarification.

\AR We acknowledge that the interpretation of $X^r$ as a "scaling dimension" requires clearer explanation. We have added the following clarifying paragraph at the end of section 7B ``Gauging Translations'':

\emph{The interpretation of $X^r$ as a "scaling dimension" follows directly from the transformation properties of the enlarged coordinate system. In equations (7.5) and (7.6), we introduce $X^r$ as an auxiliary coordinate needed to close the Poincaré algebra, initially fixing $X^r = 1$. However, when $X^r$ is allowed to vary from this reference value, the transformation properties show that the other coordinates $X^\mu$ undergo scale transformations. Specifically, in the enlarged coordinate system, physical coordinates can be interpreted as $x^\mu = X^\mu / X^r$, so variations in $X^r$ directly correspond to scaling transformations of the physical coordinates. This provides a concrete mathematical justification for interpreting $X^r$ as a scaling dimension - it is precisely the coordinate that controls the scale of the embedding. This geometric scaling property is what allows our construction to naturally accommodate the holographic principle, where the extra dimension encodes information about scales in the lower-dimensional theory.}

@@

\subsection{Section 7 - Part (b): AdS/CFT Connection}

\RC The AdS/CFT connection (Sec. 7.2) lacks concrete evidence, how does this classical construction relate to the original Maldacena's duality?

\AR We have clarified this point by adding the following text at the end of Section 7C, right after `` ... for describing the background geometry and its coupling to the string worldsheet.''

\emph{The key observation which enables this is that $X^r$ behaves as a scaling or ``holographic'' dimension as explained at the end of Sec 7B.}

\emph{Our classical construction provides a geometric framework that naturally incorporates key features of the AdS/CFT correspondence. The scaling dimension $X^r$ introduced through the mathematical necessity of closing the Poincaré algebra plays a role analogous to the radial coordinate in AdS space. Just as in AdS/CFT where the radial direction encodes the energy scale of the boundary theory, variations in our $X^r$ coordinate directly control the scale transformations of the bulk coordinates $X^\mu$. This creates a natural holographic setup where information about different energy scales in the ``boundary'' ($X^r = \text{constant}$ surfaces) is encoded in the ``bulk'' (the extended geometry with the extra $X^r$ dimension).}

\emph{However, our construction differs from standard AdS/CFT in several important ways: (1) it emerges at the classical level from gauge theory requirements rather than from quantum/non-perturbative effects, (2) the holographic direction arises from the string action itself rather than being an external bulk geometry, and (3) the correspondence is between worldsheet dynamics and the extended embedding space rather than between a boundary CFT and bulk gravity. While this provides a concrete classical realization of holographic principles, establishing a precise correspondence with Maldacena's duality would require demonstrating how our construction reproduces the specific correlation functions and scaling behaviors characteristic of AdS/CFT, which remains an open challenge for future work.}

### 2025-07-05 lqg-strings.tex
feat(T2): Complete Section 6-7 reviewer responses - Afshordi limitations & holographic clarification
- Acknowledge lack of rigorous LQG derivation for Afshordi constraints, extend caveats to (2+1)D speculation
- Clarify X^r scaling dimension via transformation properties, connect AdS/CFT with appropriate limitations
- Progress: 9/13 steps complete, ready for Discussion improvements
- Updated memory bank

@@

If one wants to treat the $X$'s as elements of a vector which transforms under both translations and rotations as in \eqref{eqn:poincare-symmetries}, we must extend the number of spacetime dimensions by one, introducing an ``auxiliary'' dimension whose value is initially set to be $X^r = 1$. Thus we have:
\begin{equation}
	\begin{pmatrix}
		X^\mu \\
		1
	\end{pmatrix} \rarrow \begin{pmatrix}
		X'^\mu \\
		1
	\end{pmatrix} = \begin{pmatrix}
		\Lambda^\mu{}_\mu & \delta X^\mu \\
		0                 & 1
	\end{pmatrix} \begin{pmatrix}
		X^\mu \\
		1
	\end{pmatrix}
\end{equation}
Now, one might be inclined to treat the new co-ordinate $X^r$ as purely an auxiliary variable which will not affect the physics. However, if we stare at the expression $X^r = 1$ for a moment, we will realise that this expression is the same form as the Dirichlet boundary condition for the end-points of a string. And, from what we know about strings with Dirichlet boundary conditions is that the constraint surface on which the end-points live will end up behaving like a dynamical object or ``D-brane'' \cite{Dai1989New-Connections,Polchinski1995Dirichlet-Branes,Horava1989Background}. In our case we now have a new enlarged bulk $\mb{M}$ with $d+1$ space dimensions - treating the auxiliary variable as a co-ordinate which can vary - and one time dimensions, or $d+2$ dimensions in total. If we fix $X^r = 1$, that leaves us with a $d+1$ submanifold $\mb{\Gamma} \in \mb{M}$ on which the string end-point is free to move.

@@

@@

>>>>

Our constraint equation (\ref{eqn:X-masses-constraint}) represents an attempt to bridge Afshordi's phenomenological emergent dimension model with LQG's general prediction of modified physics at the Planck scale. However, the specific form of these constraints and their derivation from first-principles LQG remains an open problem. While both frameworks involve discreteness - Afshordi's through hierarchical mass parameters and LQG through quantized geometric operators - establishing a concrete mathematical connection requires significantly more work.

This represents an important limitation of our current approach and highlights the need for future research to either derive such constraints from canonical LQG or develop alternative frameworks that more naturally incorporate both emergent dimensions and quantum geometric discreteness. Similar caveats also apply to our suggestion that \eqref{eqn:X-masses-constraint} should imply that spacetime is $(2 + 1)$-dimensional.

@@

If one wants to treat the $X$'s as elements of a vector which transforms under both translations and rotations as in \eqref{eqn:poincare-symmetries}, we must extend the number of spacetime dimensions by one, introducing an ``auxiliary'' dimension whose value is initially set to be $X^r = 1$. Thus we have:
\begin{equation}\label{eqn:enlarged-Xs-transform}
	\begin{pmatrix}
		X^\mu \\
		1
	\end{pmatrix} \rarrow \begin{pmatrix}
		X'^\mu \\
		1
	\end{pmatrix} = \begin{pmatrix}
		\Lambda^\mu{}_\mu & \delta X^\mu \\
		0                 & 1
	\end{pmatrix} \begin{pmatrix}
		X^\mu \\
		1
	\end{pmatrix}
\end{equation}
Now, one might be inclined to treat the new co-ordinate $X^r$ as purely an auxiliary variable which will not affect the physics. However, if we stare at the expression $X^r = 1$ for a moment, we will realise that this expression is the same form as the Dirichlet boundary condition for the end-points of a string. And, from what we know about strings with Dirichlet boundary conditions is that the constraint surface on which the end-points live will end up behaving like a dynamical object or ``D-brane'' \cite{Dai1989New-Connections,Polchinski1995Dirichlet-Branes,Horava1989Background}. In our case we now have a new enlarged bulk $\mb{M}$ with $d+1$ space dimensions - treating the auxiliary variable as a co-ordinate which can vary - and one time dimensions, or $d+2$ dimensions in total. If we fix $X^r = 1$, that leaves us with a $d+1$ submanifold $\mb{\Gamma} \in \mb{M}$ on which the string end-point is free to move.

@@

Let us note that $X^r$ can be naturally interpreted as a ``scaling dimension'' following directly from the transformation properties of the enlarged coordinate system. In equations \eqref{eqn:enlarged-Xs} and \eqref{eqn:enlarged-Xs-transform}, we introduce $X^r$ as an auxiliary coordinate needed to close the Poincaré algebra, initially fixing $X^r = 1$. However, when $X^r$ is allowed to vary from this reference value, the transformation properties show that the other coordinates $X^\mu$ undergo scale transformations. Specifically, in the enlarged coordinate system, physical coordinates can be interpreted as $x^\mu = X^\mu / X^r$, so variations in $X^r$ directly correspond to scaling transformations of the physical coordinates. This provides a concrete mathematical justification for interpreting $X^r$ as a scaling dimension - it is precisely the coordinate that controls the scale of the embedding. This geometric scaling property is what allows our construction to naturally accommodate the holographic principle, where the extra dimension encodes information about scales in the lower-dimensional theory.

@@

The key observation which enables this is that $X^r$ behaves as a scaling or ``holographic'' dimension as explained at the end of \autoref{sec:translation_gauge}.

Our classical construction provides a geometric framework that naturally incorporates key features of the AdS/CFT correspondence. The scaling dimension $X^r$ introduced through the mathematical necessity of closing the Poincaré algebra plays a role analogous to the radial coordinate in AdS space. Just as in AdS/CFT where the radial direction encodes the energy scale of the boundary theory, variations in our $X^r$ coordinate directly control the scale transformations of the bulk coordinates $X^\mu$. This creates a natural holographic setup where information about different energy scales in the ``boundary'' ($X^r = \text{constant}$ surfaces) is encoded in the ``bulk'' (the extended geometry with the extra $X^r$ dimension).

However, our construction differs from standard AdS/CFT in several important ways: (1) it emerges at the classical level from gauge theory requirements rather than from quantum/non-perturbative effects, (2) the holographic direction arises from the string action itself rather than being an external bulk geometry, and (3) the correspondence is between worldsheet dynamics and the extended embedding space rather than between a boundary CFT and bulk gravity. While this provides a concrete classical realization of holographic principles, establishing a precise correspondence with Maldacena's duality would require demonstrating how our construction reproduces the specific correlation functions and scaling behaviors characteristic of AdS/CFT, which remains an open challenge for future work.
